---
title: "Preprocess all Data"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: yes
    theme: lumen
    toc: yes
    lof: true
    number_sections: yes
biblio-style: apalike
link-citations: yes
---

```{r libraries, include=FALSE}
library(raster)
library(dplyr)
library(sf)
library(mapview)
library(magrittr)
```
# Goals

- The Goal of this script is to have TWO Dataframes at the end
  + one for all data in 50m resolution
  + one for all data in 100m resolution
- At each step we evaluated the best option to process the data (QGIS, Python, R, SNAP)

# Mask CHM with FNF
- The chm at hand has many non-noData values over man-made structures
- We only want pixels over vegetation
- For that we use the PALSAR-1 FNF Mask

```{r setpath, results="hide"}
fnf_path = "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/FNF/TCD_2015_020m_eu_03035_d05_E40N30/TCD_2015_020m_eu_03035_d05_E40N30.tif"
chm_path = "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/LAS_diff_Roda_10m_median_NA.tif"
bs_dir = "/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/final_GRDs_SNAP/"
```

## CRSs

-  read both and check for equal CRS
```{r, out.width="30%"}
# read both
fnf = raster(fnf_path)
chm = raster(chm_path)
# check for crs
crs(fnf)
crs(chm)
```

- The CRSs are not equal, so reproject the the fnf (maybe too big??)
- Definitly too big: Reproject the CHM 

```{r reprojectcmh, echo=FALSE}
chm = raster::projectRaster(chm, crs = crs(fnf)) 
```

- Now check if CRSs are equal

```{r checkifequal}
raster::compareCRS(chm, fnf)
```
## CROP FNF to CHM

```{r cropfnftochm, echo=F}
# compute shape for chm
chm_shp = st_bbox(chm) %>% st_as_sfc(., crs=crs(chm)) %>% st_sf(.)
# crop fnf
fnf_crop = crop(fnf, chm_shp)
fnf_masked = mask(fnf_crop, chm_shp)
```

- Resample the chm (10 x 10m) to the fnf (20 x 20m)

```{r resample chm, echo=F}
chm = resample(chm, fnf_masked)
```

- Set all Pixels in CHM to NA that are 0 in fnf

```{r nachm, echo=F}
chm_fnf = chm
chm_fnf[fnf_masked == 0] = NA
# check the numer of pixels with NA before and after masking with fnf
sum(is.na(values(chm)))
sum(is.na(values(chm_fnf)))
```

- Set all values below 0 in chm to NA

```{r chmnodata, echo=F}
chm_fnf[chm_fnf < 0] = NA
sum(is.na(values(chm_fnf)))
```
- reproject the chm to 32632
- and write it
```{r writechm}
chm_fnf_repro = raster::projectRaster(chm_fnf, crs = crs(raster(chm_path)))
writeRaster(chm_fnf_repro, "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_20m_maskedtofnf_32632.tif", overwrite=T)
```

- read the chm and resample it to 50m

```{r readnewchm, echo=F}
chm_new_path  = "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_20m_maskedtofnf_32632.tif"
chm_50m_path = "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_50m_maskedtofnf_32632.tif"
chm_100m_path = "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_100m_maskedtofnf_32632.tif"
cmd = sprintf("gdalwarp -tr 50 50 %s %s", chm_new_path, chm_50m_path)
cmd2 = sprintf("gdalwarp -tr 100 100 %s %s", chm_new_path, chm_100m_path)
system(cmd)
system(cmd2)
```


# Backscatter

- read all backscatters and new CHM

```{r readbackscatter}
# read chm
chm = raster("/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_20m_maskedtofnf_32632.tif")
# locaetion on disk
backscatter_paths = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/final_GRDs_SNAP", full.names = T)
backscatter_paths = backscatter_paths %>% stringr::str_subset(., pattern = "Cal_TC.tif", negate = F)
# load into list
backscatter_raster = lapply(backscatter_paths, function(x) raster(x))
# make names
names_backscatter = lapply(backscatter_raster, function(x) paste0("bs_", substr(names(x), start=40, stop=45)))
# apply names
names(backscatter_raster) = names_backscatter
```

- reproject all backscatter to 32632 and set the Source-NoData-Value to 0

```{r reprojectbackscatter}
for (i in seq_along(backscatter_paths)){
  path = "/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/32632_NoData0_GRDS"
  if(!dir.exists(path)) dir.create(path)
  base_name = strsplit(backscatter_paths[[i]], .Platform$file.sep)[[1]] %>% tail(., n=1) %>% substr(., 40,45)
  ending = ".tif"
  file_name = paste(path, paste0(base_name, "_32632", ending), sep = .Platform$file.sep)
  cmd = sprintf("gdalwarp -t_srs EPSG:32632 -srcnodata 0 %s %s", backscatter_paths[[i]], file_name)
  system(cmd)
}
```
- load backscatter back in

```{r loadbackscatter}
# naming convention
backscatter_32632_paths = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/32632_NoData0_GRDS", full.names = T)
backscatter_32632_raster = lapply(backscatter_32632_paths, function(x) raster(x))
names_backscatter = lapply(backscatter_32632_paths, function(x) paste0("bs_", substr(x, start=72, stop=77)))
names(backscatter_32632_raster) = names_backscatter
```

- compute intersection between one backscatter and the chm

```{r}
# bs_sf = st_bbox(backscatter_32632_raster[[1]]) %>% st_as_sfc(., crs=32632)
# chm_sf = st_bbox(chm) %>% st_as_sfc(., crs=32632)
# bs_sf = st_transform(bs_sf, 32632)
# chm_sf = st_transform(chm_sf, 32632)
# # Calculate Intersection
# intersec = st_intersection(bs_sf, chm_sf)
# ints = intersec %>% st_sf()
```


- gdalwarp all backscatters to (50x50 and 100x100)
```{r gdalwarpbs}
# get all backscatter
backscatter_paths = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/32632_NoData0_GRDS/", full.names = T)
resolutions = c("50 50", "100 100")
for (i in resolutions){
  for (j in seq_along(backscatter_paths)){
    name_new = paste0(strsplit(backscatter_paths[[j]], .Platform$file.sep)[[1]] %>%  tail(., n=1) %>% substr(., start = 1, stop = nchar(.)-4), "_32632", "_", substr(i,1,2), "m", ".tif")
    dir_new = "/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/backscatter_resampled_50_100m"
    if(!dir.exists(dir_new)) dir.create(dir_new)
    name_new = paste(dir_new, name_new, sep = .Platform$file.sep)
    cmd = sprintf("gdalwarp -r med -tr %s  %s %s", i, backscatter_paths[[j]], name_new)
    print(cmd)
    system(cmd)
    
  }
}
```

***
***
***

END OF FIRST PREPROCESSING

- load 50 and 100m chm
```{r, eval=F}
chm_50 = raster(chm_50m_path)
chm_100 = raster(chm_100m_path)
```

- load 50m backscatter
```{r, eval=F}
backscatter_50m_paths = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/final_GRDs_R/", full.names = T)
backscatter_50m_paths = grep("_50m.tif", backscatter_50m_paths, value = T)
backscatter_50m_raster = lapply(backscatter_50m_paths, function(x) raster(x))
backscatter_100m_paths = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/final_GRDs_R/", full.names = T)
backscatter_100m_paths = grep("_10m.tif", backscatter_100m_paths, value = T)
backscatter_100m_raster = lapply(backscatter_100m_paths, function(x) raster(x))
```


- crop and mask all backscatter to intersection
```{r, eval=F}
backscatter_32632_raster_cropped = lapply(backscatter_32632_raster, function(x) crop(x, ints))
backscatter_32632_raster_masked = lapply(backscatter_32632_raster_cropped, function(x) crop(x, ints))
# write them
new_dir = "/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/maskedBackscatter"
if(!dir.exists(new_dir)) dir.create(new_dir)
for (i in seq_along(backscatter_32632_raster_masked)){
  n = paste0(paste(new_dir, names(backscatter_32632_raster_masked)[[i]], sep = .Platform$file.sep), ".tif")
  writeRaster(backscatter_32632_raster_masked[[i]], n)
}
```

- crop and mask chm to intersection
```{r, eval=F}
chm_cropped_50 = crop(chm_50, ints)
chm_masked_50 = mask(chm_cropped_50, ints)
chm_cropped_100 = crop(chm_100, ints)
chm_masked_100 = mask(chm_cropped_100, ints)
# write it
writeRaster(chm_masked_50, "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_masked_to_intersection_50m.tif")
writeRaster(chm_masked_100, "/home/robin/geodata/geo411/GEO411_FSH_Roda/ancillary_data/chm_masked_to_intersection_100m.tif")
```


- crop and mask backscatter to intersection

```{r, eval=F}
backscatter_100m_cropped_intersection = lapply(backscatter_100m_raster, function(x) crop(x, ints))
backscatter_100m_masked_intersection = lapply(backscatter_100m_cropped_intersection, function(x) mask(x, ints))
backscatter_50m_cropped_intersection = lapply(backscatter_50m_raster, function(x) crop(x, ints))
backscatter_50m_masked_intersection = lapply(backscatter_50m_cropped_intersection, function(x) mask(x, ints))
```

- Resample chms to backscatter to get exactly equal extent
```{r, eval=F}
chm_masked_50 = resample(chm_masked_50, backscatter_50m_masked_intersection[[1]])
chm_masked_50[chm_masked_50 < 0] = NA
chm_masked_100 = resample(chm_masked_100, backscatter_100m_masked_intersection[[1]])
chm_masked_100[chm_masked_100 < 0] = NA
```

- write Data

```{r, eval=F}
# write 50m Backscatter
for (i in seq_along(backscatter_50m_masked_intersection)){
  name = names(backscatter_50m_masked_intersection[[i]])
  new_dir = "/home/robin/geodata/geo411/GEO411_FSH_Roda/snap_grd/masked_resampled_32632_backscatter"
  if(!dir.exists(new_dir)) dir.create(new_dir)
  name_final = paste(new_dir, paste0(name, ".tif"), sep=.Platform$file.sep)
  writeRaster(backscatter_50m_masked_intersection[[i]], name_final)
}
```

# Coherence

- find Coherence

```{r}
files = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda", pattern = "topophase.cor.geo$", full.names = T, recursive = T)
files = grep("HH|test", files, invert = T, value = T)
```

- gdal_translate coherence to make tiff
```{r}
new_dir = "/home/robin/geodata/geo411/GEO411_FSH_Roda/coherences"
if(!dir.exists(new_dir)) dir.create(new_dir)
dates = sapply(files, function(x) substr(x, 44, 60))
files_new = paste(new_dir, paste0(dates, ".tif"), sep = .Platform$file.sep)
# translate them
for(i in seq_along(files)){
  cmd = sprintf("gdal_translate -b 2 -a_nodata 0 %s %s", files[[i]], files_new[[i]])
  system(cmd)
}
```


- gdalwarp them to 32632 and 50x50 and 100x100

```{r}
resolutions = c("50 50", "100 100")
files = list.files("/home/robin/geodata/geo411/GEO411_FSH_Roda/coherences", full.names = T)
for (i in seq_along(resolutions)){
  for (j in seq_along(files)){
    base_path =  strsplit(files[[j]], split = .Platform$file.sep)
    base_path = paste(base_path[[1]][1:lengths(base_path)-1], collapse = .Platform$file.sep)
    date = substr(files[[j]], 55, 71)
    file_path = paste(base_path, paste0(date, "_32632_", substr(resolutions[[i]],1,2), ".tif"), sep = .Platform$file.sep)
    cmd = sprintf("gdalwarp -tr %s -r med -t_srs EPSG:32632 %s %s", resolutions[[i]], files[[j]], file_path)
    system(cmd)
}}
```

